Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}

### 2023-05-18 Papers/Paper CASC format/paper.tex
finished paper for CASC

 \newpage 
%just for now 
\section{Parametric Groebner Systems}
\label{pgs}

% This section covers key concepts for studying parametric 
% polynomial ideals, including parametric rings and specialization 
% ideals and comprehensive Gr{\"o}bner systems for parametric polynomials that generalize 
% Gr{\'o}bner basis in section \ref{4sec1}. Section 2 explains why comprehensive Gr{\"o}bner systems are advantageous 
% and some problems that arise without CGS. Section 3 
% extends the ideal membership problem to parametric ideals.
In this section, we delve into essential concepts for 
parametric polynomial ideals. These include parametric rings, 
specialized ideals, and comprehensive Gr{\"o}bner systems, which are built upon 
Gr{\"o}bner basis theory. 
We discuss the benefits of using comprehensive Gr{\"o}bner systems 
and how they help address problems that occur when not using them. 
Furthermore, we expand on the ideal membership problem to encompass parametric ideals.

% \subsection{Parametric Polynomials and Comprehensive Groebner Systems}
% \label{4sec1}

% This section introduces the general setting of parametric polynomial rings and a ring homomorphism that 
% represents the choice of parameters. Moreover, we define comprehensive Gr{\"o}bner systems (CGS),
% and show some examples of (CGS) using our implementation of a parametric Buchberger's algorithm.

\begin{definition}
    Let $\mathcal{K}$ be a field. The ring of parametric polynomials
    $\mathcal{K}[a_1,..,a_s][x_1,...,x_n]$ is a ring of polynomials in the variables $x_1,...,x_n$ with coefficients that are polynomials
    in the ring $\mathcal{K}[a_1,..,a_s]$. 
    For simplicity, we denote this ring by $\mathcal{K}[A][X]$.
\end{definition}

@@

\end{definition}
% For example let $P=a x+by \in \Rr[a,b][x,y]$. 
% a specialization $\sigma$ is defined by $P\rightarrow P(2,5)$ so that $$\sigma (P)=\sigma(a x+by)$$
% and since $\sigma $ is a ring homomorphism we have,
% $$\sigma(a x+by)= \sigma(a) \sigma(x)+\sigma(b)\sigma(y)$$
% $$= 2x+5y.$$
It is easier to think about specializations as a choice of parameters. \\

Comprehensive Gr{\"o}bner systems extend Gr{\"o}bner basis to 
parametric polynomials and have diverse practical applications in 
algebraic geometry, combinatorics, coding theory, computer-aided design, and robotics.
\begin{definition}
    Let $\Ii$ be a finitely generated, non-zero ideal in the ring $\mathcal{K}[A][X]$ with a fixed monomial order $\prec$ on the variables. 
    Let $\{A_1,\dots,A_k\}$ be a family of locally closed sets that forms a partition of $L^m$ 
    , such that the disjoint union $\bigcup\limits_{i=1}^k A_i=L^m$, and $\{G_1,\dots,G_k\}$ 
    be a family of subsets in $\mathcal{K}[A][X]$.
    The system, 
	\begin{equation}\label{CGS}
	\left\{
		\begin{array}{ll}
			G_1 & A_1 \\
			\vdots  & \vdots \\ 
			G_k & A_k \\
		\end{array}
		\right. 
	\end{equation}

    is called a comprehensive Gr{\"o}bner system (CGS) of $I$
     if for all specializations $\sigma_i \in A_i, \sigma_i (G_i)$ 
     is a Gr{\"o}bner basis of the specialized ideal $\sigma_i (I)\subset \mathcal{K}[X]$.
     Where each pair $\{G_i,A_i\}$ is called a branch. \cite{Kapur}
\end{definition}
Here, the notation $\sigma_i (\Ff)$, with $\Ff=\{f_1,f_2,\dots ,f_k\}$, is the specialization $\sigma_i$
of each $f_j$ such that  $\sigma_i (\Ff)=\{\sigma_i(f_j)~|~ f_j\in \Ff\}$
and  $"\sigma_i \in A_i"$ stands for a specialization given by $P\rightarrow P(a_{1_0},\dots,a_{m_0})$ 
for $(a_{1_0},\dots,a_{m_0})\in A_i$.\\

In this paper, we choose $A_i$ to be of the form 
$$V(p_1,\dots,p_s)\backslash V(q_1,\dots,q_t) \subset L^m.$$
Where $p_1,\dots,p_s,q_1,\dots,q_t\in \mathcal{K}[A]$ and $V(p_1,\dots,p_s)$ is a Variety.

A proof of existence and the first algorithm to compute the (CGS) were introduced by Volker Weispfenning.
\begin{theorem}
    Let $\mathcal{I}\subseteq\mathcal{K}[A][X].$ The corresponding comprehensive Gr{\"o}bner system exists.\cite{weispfenning}
\end{theorem}
We use our implementation of a parametric Buchberger's algorithm to compute (CGS). 
This is illustrated in the following examples.
\begin{example}
    Let $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by 
    $$\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}.$$
    The comprehensive Gr{\"o}bner system of $\Ii$ with a Lexicographic order $y\prec x$ is:
    {\color{red} we use our code here}
	% $$
    % % \begin{cases}\left\{a^2 b^2 y^3+a b^4 y^4-2 a b^3 y^3,x-b^2
    % %     y^2\right\} &  a b \neq 0 \\
    % % \{x\} &  b=0 \\
    % % \left\{b^2 y^2-x\right\} &  a=0 \land b \neq 0.
    % % \end{cases}
    % $$
\end{example}
The idea of a comprehensive Gr{\"o}bner system is to provide a "parameter sensitive Gr{\"o}bner basis".
It branches the cases of the parameters that affect leading terms. For instance, in the previous
example, if $a=0$, then the leading term $a x^2$ vanishes which leads to a different Gr{\"o}bner basis. Therefore,
the (CGS) considers cases of the leading coefficients parameters Vanishing.

\begin{example}
    Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
    $$\Ff=\left\{a^2 x-y,-a+b^2 x
    y-y^2\right\}.$$
    The comprehensive Gr{\"o}bner system of $\mathcal{I}$ with a Lexicographic order $y\prec x$ is:

@@

Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters.
In practice, to compute the Gr{\"o}bner basis without considering the parameters, we assume non-zero parameters by default.
This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$. i.e., 
$\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialized ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
Next, we show the impact of this assumption.
\begin{enumerate}
    \item In general $\sigma\left(\Gg\right)$ is not necessarily a Gr{\"o}bner basis.
    \begin{example}
        \label{ex431}
        Let $\Ii=\langle 
           {a x^2 - y, -2 + b x y} \rangle$. The corresponding Gr{\"o}bner basis 
           is $\Gg=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$. Consider 
           the specialization $\sigma$ given by $P\rightarrow P(0,1)$.
           The set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
		   This can be shown by the Buchberger Criterion (see Chapter 2 in\cite{Cox}).

        %    To show that we find the S-polynomials of each pair in the set and we find that: $S(-y^2, -2 + x y)=-2 y$.
        %    We divide $S(-y^2, -2 + x y)$ by $\sigma(\Gg)$ and find that the remainder is $-2y$. 
        %    Therefore by the Buchberger Criterion, $\sigma(\Gg)$ is not a Gr{\"o}bner basis.
    \end{example}
    \item Consequently, the algorithm utilized to resolve the ideal membership 
    problem in chapter 2 of \cite{Cox} can not be used for all specializations.

    \begin{example}
        \label{ex433}
        Consider the ideal $\Ii$ in the above example. The corresponding Gr{\"o}bner basis computed under the assumption of non-zero parameters
        is $$\sigma_0(\Gg)=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$$ where $\sigma_0 \in \Rr^2\backslash V(a,b)$. 
        Now, it is possible to check if a polynomial $f\in \Rr[x,y]$ belongs to $\sigma(\Ii)$ only for $\sigma_0 $     
        because we know that $\sigma_0(\Gg)$ is a Gr{\"o}bner basis. For other specializations such as 
        the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as shown in the above example. 
        Therefore, we can not proceed with the division algorithm as we can not guarantee a unique remainder if we use $\sigma(\Gg)$.  
    \end{example}

\end{enumerate}

@@

{\color{red} Is this a definition?}
 Consider, $f\in \Kk[A][X]$ we say that $f\in\Ii$ under $\sigma$, if there exists a specialization 
 $\sigma\in L^m$ such that $\sigma(f)\in\sigma(\Ii)$.\\

To determine if $f\in\Ii$ we consider the (CGS) of $\Ii$ given \ref{CGS}. 
We check if $f\in\sigma_i(\Ii)$ for each $i\in\{1,..,k\}$ and $\sigma_i\in A_i$.
For each $\sigma_i(\Ii)$ we have the corresponding Gr{\"o}bner basis $\sigma_i(G_i)$
so we divide $f$ by the $\sigma_i(G_i)$ if the remainder is zero for some $i\in\{1,..,k\}$,
then $f\in\sigma_i(\Ii)$. Therefore, $f\in\Ii$. However, if the remainder is non-zero for all $i$ we conclude that 
$f\notin\Ii$.

\begin{example}
	Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
    $$\Ff=\{a^2 x - y, -b^2 x y^2 - x\}.$$
	We will determine if $f(x,y)=-x+y^2\in\mathcal{I}$.
	We compute the corresponding (CGS)
	{\color{red} add example 3}\\
	First let $A_1=\{(a,b)~|~ a\ne 0, b\ne 0\}$ and $A_2=\{(a,b)~|~ ab=0\}$. 
	We divide $f$ by $\{\frac{y}{b^2}+y^3,x-\frac{y}{a^2}\}$ and we obtain the remainder $-\frac{y}{a^2}+y^2$ and 
	since the exists no specialization in $A_1$ such that this remainder is zero we conclude that $f\notin \sigma_1(\Ii)$ for
	$\sigma_1\in A_1$. Now we repeat the reduction of $f$ but with the second Gr{\"o}bner basis $\{x,y\}$. We obtain a zero remainder, which implies that 
	$f\in \sigma_2(\Ii)$ for $\sigma_2\in A_2$. Therefore, $f\in\Ii$ under a specialization $\sigma_2\in A_2$.
\end{example}

\newpage
\section{The time-explicit case}
\label{time-exp}
In this section, we introduce Algorithm \ref{alg: discreteconservedqexplicit}  for checking the conservation of a quantity for time-explicit schemes. We discuss the algorithm for general schemes in the next section.
\begin{algorithm}
	\DontPrintSemicolon
	\KwIn{$\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{D}(\Nn))$ and $\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ right-hand side of a time-explicit scheme}
	\KwOut{ {\sc True}, if $\mathcal{J}$ is conserved in time, {\sc False} otherwise }

@@

	\caption{{\sc time-explicit DiscreteConservedQ}}
	\label{alg: discreteconservedqexplicit}
\end{algorithm}
The result of Step 2 exists only at time $t$ because all instances of $t+1$ have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
\begin{example}
	\label{timeexplicitheateqalgo}
	We examine the conservation of $\sum_nu(n,t)$ under the following scheme.
	\begin{equation*}
		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t)).
	\end{equation*}
	The discrete time derivative (Step 1) is
	\begin{gather*}
		\sum_{n\in\Ii}u(n,t+1)-u(n,t).
	\end{gather*}
	Replace $u(n,t+1)$ by the right-hand side (Step 2) to get
	\begin{gather*}
		\sum_{n\in\Ii}u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t))-u(n,t).
	\end{gather*}
	Then, we compute the discrete partial variational derivative (Step 3), which equals zero. Hence (Step 4), we have conservation.
\end{example}
\begin{example}
	\label{heatex1}
	We verify the result from Example \ref{timeexplicitheateqalgo}, using our implementation in {\sc Mathematica} of Algorithm \ref{alg: discreteconservedqexplicit}.
	\begin{EXE}
		LABEL="rhs1"
		(*CODE*)
		variables = Association["indVars"->{n},"depVars"->{u},"rhs"->{u[n]+(u[n+1]-2*u[n]+u[n-1])}]
		expression = u[n]
		DiscreteConservedQOperator[variables][expression]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/rhs1ctt.ed.tex}
		\input{executables/rhs1ctt.result.ed.tex}
	\end{small}
\end{example}

@@

	The algorithm treats the discrete time derivative $\sum_{n\in\Ii}u(n+1,t+1)-u(n+1,t)$
	as the polynomial $z_1-z_2$.

@@

	\begin{equation*}

@@

	\end{equation*}

@@

\noindent{\bf Step 3 and 4: Compute Gr{\"o}bner basis and reduce the discrete time derivative.}\\
\emph{Polynomial ideals and Gr{\"o}bner bases.}\label{polid}
In the previous step, we identified a finite set of polynomials comprising the translated scheme ${H^1,\ldots,H^r}$ and the discrete time derivative, each containing a finite number of instances of $u_j(n+e,t+l)$. We now proceed to reduce $\mathcal{T}$ by treating it as an element of a polynomial ring and performing multivariate polynomial division with respect to the translated scheme. To accomplish this, we adapt the definitions and theorems presented in \cite{hibi13} to suit our specific context.
\begin{definition}
	Let $\mathcal{K}$ be as in Definition \ref{diffideal}. Let $\mathcal{R}$ be the polynomial ring generated by $\mathcal{K}$ and all instances of $u_j(n+e,t+l)$ that occur in $\{H^1,\ldots,H^r\}$.
	We call a set $ I\subseteq \mathcal{R}$ a \emph{(polynomial) ideal} if
	$p_1,p_2\in I$ implies $p_1+p_2\in I$ and
	$p_1\in I,p_2\in\mathcal{R}$ implies $p_1p_2\in I$.
\end{definition}
\begin{definition}
	We denote by $\langle H^1,\ldots,H^r\rangle$ the smallest (polynomial) ideal containing $\{H^1,\ldots,H^r\}$.
\end{definition}
Given the ideal $\langle H^1,\ldots,H^r\rangle$ and the discrete time derivative $\mathcal{T}\in\mathcal{R}$, we want to determine if $\mathcal{T}\in\langle H^1,\ldots,H^r\rangle$ or if we can write $\mathcal{T}$ in a simpler form, using $\{H^1,\ldots,H^r\}$.
Hence, using multivariate polynomial division, we search for $p^1,\ldots,p^r,p^{r+1}\in\mathcal{R}$ such that
\begin{gather*}
	\mathcal{T}=p^1H^1+\ldots+p^rH^r+p^{r+1}.
\end{gather*}
Unfortunately, the resulting remainder $p^{r+1}$ may not be unique \cite[page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although $\mathcal{T}$ belongs to $\langle H^1,\ldots,H^r \rangle$. 
Replacing $\{H^1,\ldots,H^r\}$ by a Gr{\"o}bner basis for the ideal $\langle H^1,\ldots,H^r \rangle$ guarantees the uniqueness of the remainder of the polynomial division. 
Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, a Gr{\"o}bner basis for the polynomial ideal always exists and is finite.
A Gr{\"o}bner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.

@@

	\caption{{\sc FindDiscreteConservedQuantityBasis}}
	\label{alg: findconservedquantitybasis}
\end{algorithm}
We have implemented this algorithm in {\sc FindDiscreteConservedQuantityBasis}.
\begin{example}
In the context of the heat equation from Example \ref{heatex1}, we aim to find a basis for conserved quantities generated by $u(n,t)$ and $n$ with a maximum degree of 3.
	\begin{EXE}
		LABEL="heateq11"
		(*CODE*)
		variables = <|"indVars"->{n},"depVars"->{u},"rhs"->{u[n]+u[n+1]-2*u[n]+u[n-1]}|>
		FindDiscreteConservedQuantityBasisOperator[variables][<|"degree"->3,"generators"->{u[n],n}|>]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/heateq11ctt.ed.tex}
		\input{executables/heateq11ctt.result.ed.tex}
	\end{small}
The first conserved quantity corresponds to mass conservation and the second to center of mass conservation. 
\end{example}
\section{Examples and applications}
\label{eaa}
We demonstrate our code's capabilities using two examples: the inviscid Burgers equation and a mean-field game system. 
Through these examples, we examine different discretizations, evaluate their conservation properties, and identify conserved quantities.
Although the examples presented feature only first-order time derivatives, the algorithm can handle PDEs with derivatives of any order.

@@

		v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,\quad

@@

	The related backward-forward system reads
		-v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,\quad

@@

@@

\subsection{Concluding remarks}

>>>>

 %\newpage 
%just for now 
\section{Parametric Groebner Systems}
\label{pgs}

% This section covers key concepts for studying parametric 
% polynomial ideals, including parametric rings and specialization 
% ideals and comprehensive Gr{\"o}bner systems for parametric polynomials that generalize 
% Gr{\'o}bner basis in section \ref{4sec1}. Section 2 explains why comprehensive Gr{\"o}bner systems are advantageous 
% and some problems that arise without CGS. Section 3 
% extends the ideal membership problem to parametric ideals.
Here, we review essential concepts for 
parametric polynomial ideals. These include parametric rings, 
specialized ideals, and comprehensive Gr{\"o}bner systems, which are built upon 
Gr{\"o}bner basis theory. 
We discuss the advantages of comprehensive Gr{\"o}bner systems 
compared with standard Gr\"obner basis. 
Furthermore, we expand on the ideal membership problem to encompass parametric ideals.

\begin{definition}
The ring of parametric polynomials
    $\mathcal{K}[a_1,..,a_s][x_1,...,x_n]$ in a field $\mathcal{K}$ is a ring of polynomials in the variables $x_1,...,x_n$ with coefficients that are polynomials
    in the ring $\mathcal{K}[a_1,..,a_s]$. 
    To simplify, we denote this ring by $\mathcal{K}[A][X]$.
\end{definition}

@@

\end{definition}
It is easier to think about specializations as a choice of parameters. 
Comprehensive Gr{\"o}bner systems extend Gr{\"o}bner basis to 
parametric polynomials.
% and have diverse applications in 
%algebraic geometry, combinatorics, coding theory, computer-aided design, and robotics.
\begin{definition}
    Let $\Ii$ be a finitely generated, non-zero ideal in the ring $\mathcal{K}[A][X]$ with a fixed monomial order $\prec$ on the variables. 
    Let $\{A_1,\dots,A_k\}$ be a family of locally closed sets that forms a partition of $L^m$ 
    , such that the disjoint union $\bigcup\limits_{i=1}^k A_i=L^m$, and $\{G_1,\dots,G_k\}$ 
    be a family of subsets in $\mathcal{K}[A][X]$.
    The system, 
	\begin{equation}\label{CGS}
	\left\{
		\begin{array}{lll}
			G_1 &\quad& A_1 \\
			\vdots  &\quad& \vdots \\ 
			G_k &\quad & A_k \\
		\end{array}
		\right. 
	\end{equation}
    is a {\it comprehensive Gr{\"o}bner system} (CGS) of $I$
     if for all specializations $\sigma_i \in A_i, \sigma_i (G_i)$ 
     is a Gr{\"o}bner basis of the specialized ideal $\sigma_i (I)\subset \mathcal{K}[X]$.
     Where each pair $\{G_i,A_i\}$ is called a branch. \cite{Kapur}
\end{definition}
Here, the notation $\sigma_i (\Ff)$, with $\Ff=\{f_1,f_2,\dots ,f_k\}$, is the specialization $\sigma_i$
of each $f_j$ such that  $\sigma_i (\Ff)=\{\sigma_i(f_j)~|~ f_j\in \Ff\}$
and  $"\sigma_i \in A_i"$ stands for a specialization given by $P\rightarrow P(a_{1_0},\dots,a_{m_0})$ 
for $(a_{1_0},\dots,a_{m_0})\in A_i$.
In this paper, $A_i$ is of the form 
$V(p_1,\dots,p_s)\backslash V(q_1,\dots,q_t) \subset L^m$.
Where $p_1,\dots,p_s,q_1,\dots,q_t\in \mathcal{K}[A]$ and $V(p_1,\dots,p_s)$ is a variety.

A proof of existence and an algorithm to compute CGS appeared first in \cite{weispfenning}.
\begin{theorem}[\cite{weispfenning}]
    Let $\mathcal{I}\subseteq\mathcal{K}[A][X]$. Then, there exists a CGS . 
\end{theorem}
Here, to compute CGS, we use a parametric implementation of 
Buchberger's algorithm which is illustrated in the following examples.

{\color{red} I think we need to explain briefly the parametric implementation. 
Basically we use Buchberger algorithm but branch every time we divide by a coefficient of a polynomial, considering the cases where this coefficient vanishes and when it does not. }

\begin{example}
    Let  $\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}$, 
    and     
    $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by $\Ff$. 
    The comprehensive Gr{\"o}bner system of $\Ii$ with a Lexicographic order $y\prec x$ is:
	% $$
    % % \begin{cases}\left\{a^2 b^2 y^3+a b^4 y^4-2 a b^3 y^3,x-b^2
    % %     y^2\right\} &  a b \neq 0 \\
    % % \{x\} &  b=0 \\
    % % \left\{b^2 y^2-x\right\} &  a=0 \land b \neq 0.
    % % \end{cases}
    % $$
\end{example}
The idea of a comprehensive Gr{\"o}bner system is to provide a "parameter sensitive Gr{\"o}bner basis".
It branches the cases of the parameters that affect leading terms. For instance, in the previous
example, if $a=0$, then the leading term $a x^2$ vanishes which leads to a different Gr{\"o}bner basis. Therefore,
the (CGS) considers cases of the leading coefficients parameters vanishing.

\begin{example}
	Let $\Ff=\left\{a^2 x-y,-a+b^2 x
	y-y^2\right\}$, and 
 $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by $\Ff$.
    The comprehensive Gr{\"o}bner system of $\mathcal{I}$ with a Lexicographic order $y\prec x$ is:

@@

Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters; that is, 
computing Gr{\"o}bner basis assuming the non-zero coefficients.
This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$, i.e., 
$\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialized ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
%Next, we show the impact of this assumption.
In general $\sigma\left(\Gg\right)$ is not necessarily a Gr{\"o}bner basis.
    \begin{example}
        \label{ex431}
        Let $\Ii=\langle 
           {a x^2 - y, -2 + b x y} \rangle$. The corresponding Gr{\"o}bner basis 
           is $\Gg=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$. Consider 
           the specialization $\sigma$ given by $P\rightarrow P(0,1)$.
           The set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
		   This can be shown by the Buchberger Criterion (see Chapter 2 in\cite{Cox}).

        %    To show that we find the S-polynomials of each pair in the set and we find that: $S(-y^2, -2 + x y)=-2 y$.
        %    We divide $S(-y^2, -2 + x y)$ by $\sigma(\Gg)$ and find that the remainder is $-2y$. 
        %    Therefore by the Buchberger Criterion, $\sigma(\Gg)$ is not a Gr{\"o}bner basis.
    \end{example}

Consequently, the algorithm utilized to resolve the ideal membership 
    problem in Chapter 2 of \cite{Cox} can not be used for all specializations.

    \begin{example}
        \label{ex433}
        Consider the ideal $\Ii$ in the above example. The corresponding Gr{\"o}bner basis computed under the assumption of non-zero coefficients
        is $$\sigma_0(\Gg)=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$$ where $\sigma_0 \in \Rr^2\backslash V(a,b)$. 
        Now, it is possible to check if a polynomial $f\in \Rr[x,y]$ belongs to $\sigma(\Ii)$ only for $\sigma_0 $     
        because we know that $\sigma_0(\Gg)$ is a Gr{\"o}bner basis. For other specializations such as 
        the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as shown in the above example. 
        Therefore, we can not proceed with the division algorithm as we can not guarantee a unique remainder.% if we use $\sigma(\Gg)$.  
    \end{example}

@@

{\color{red} Is this a definition? DG: I think now it is better }
Let $f\in \Kk[A][X]$. We say that $f\in\Ii$ under 
 $\sigma\in L^m$ if  $\sigma(f)\in\sigma(\Ii)$.\\

To determine  $f\in\Ii$, we consider the (CGS) of $\Ii$ given \ref{CGS}. 
We check if $f\in\sigma_i(\Ii)$ for each $i\in\{1,..,k\}$ and $\sigma_i\in A_i$.
For each $\sigma_i(\Ii)$, we have the corresponding Gr{\"o}bner basis $\sigma_i(G_i)$. Thus, 
we divide $f$ by the $\sigma_i(G_i)$. If the remainder is zero for some $i\in\{1,..,k\}$,
then $f\in\sigma_i(\Ii)$. Therefore, $f\in\Ii$. However, if the remainder is non-zero for all $i$, we conclude that 
$f\notin\Ii$.

\begin{example}
	Let $\Ff=\{a^2 x - y, -b^2 x y^2 - x\}$ and let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
    $\Ff$.
	We will determine if $f(x,y)=-x+y^2\in\mathcal{I}$.
	We compute the corresponding (CGS)
	{\color{red} add example 3}\\
	First let $A_1=\{(a,b)~|~ a\ne 0, b\ne 0\}$ and $A_2=\{(a,b)~|~ ab=0\}$. 
	We divide $f$ by $\{\frac{y}{b^2}+y^3,x-\frac{y}{a^2}\}$ and we obtain the remainder $-\frac{y}{a^2}+y^2$ and 
	since the exists no specialization in $A_1$ such that this remainder is zero we conclude that $f\notin \sigma_1(\Ii)$ for
	$\sigma_1\in A_1$. Now we repeat the reduction of $f$ but with the second Gr{\"o}bner basis $\{x,y\}$. We obtain a zero remainder, which implies that 
	$f\in \sigma_2(\Ii)$ for $\sigma_2\in A_2$. Therefore, $f\in\Ii$ under a specialization $\sigma_2\in A_2$.
\end{example}

\section{The time-explicit case}
\label{time-exp}
In this section, we introduce Algorithm \ref{alg: discreteconservedqexplicit}  for checking the conservation of a quantity for time-explicit schemes. We discuss the algorithm for general schemes in the next section.
\begin{algorithm}
	\DontPrintSemicolon
	\KwIn{$\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{D}(\Nn))$ and $\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ right-hand side of a time-explicit scheme}
	\KwOut{ {\sc True}, if $\mathcal{J}$ is conserved in time, {\sc False} otherwise }

@@

	\caption{{\sc time-explicit DiscreteConservedQ}}
	\label{alg: discreteconservedqexplicit}
\end{algorithm}
The result of Step 2 exists only at time $t$ because all instances of $t+1$ have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
\begin{example}
	\label{timeexplicitheateqalgo}
	We examine the conservation of $\sum_nu(n,t)$ under the following scheme.
	$
		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t)).
	$
	The discrete time derivative (Step 1) is
	$
		\sum_{n\in\Ii}u(n,t+1)-u(n,t).
$
	Replace $u(n,t+1)$ by the right-hand side (Step 2) to get
	$
		\sum_{n\in\Ii}u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t))-u(n,t).
	$
	Then, we compute the discrete partial variational derivative (Step 3), which equals zero. Hence (Step 4), we have conservation.
\end{example}
\begin{example}
	\label{heatex1}
	We verify the result from Example \ref{timeexplicitheateqalgo}, using our implementation in {\sc Mathematica} of Algorithm \ref{alg: discreteconservedqexplicit}.
	\begin{EXE}
		LABEL="rhs1"
		(*CODE*)
		variables = Association["indVars"->{n},"depVars"->{u},"rhs"->{u[n]+(u[n+1]-2*u[n]+u[n-1])}]
		expression = u[n]
		DiscreteConservedQOperator[variables][expression]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/rhs1ctt.ed.tex}
		\input{executables/rhs1ctt.result.ed.tex}
	\end{small}
\end{example}

@@

%\begin{example}
%	The algorithm treats the discrete time derivative $\sum_{n\in\Ii}u(n+1,t+1)-u(n+1,t)$
%	as the polynomial $z_1-z_2$.
%\end{example}

@@

@@

@@

\noindent{\bf Step 3 and 4: Compute Gr{\"o}bner basis and reduce the discrete time derivative.}\\
\emph{Polynomial ideals and Gr{\"o}bner bases.}\label{polid}
In the previous step, we identified a finite set of polynomials comprising the translated scheme ${H^1,\ldots,H^r}$ and the discrete time derivative, each containing a finite number of instances of $u_j(n+e,t+l)$. We now proceed to reduce $\mathcal{T}$ by treating it as an element of a polynomial ring and performing multivariate polynomial division with respect to the translated scheme. To accomplish this, we adapt the definitions and theorems presented in \cite{hibi13} to suit our specific context.
\begin{definition}
	Let $\mathcal{K}$ be as in Definition \ref{diffideal}. Let $\mathcal{R}$ be the polynomial ring generated by $\mathcal{K}$ and all instances of $u_j(n+e,t+l)$ that occur in $\{H^1,\ldots,H^r\}$.
	We call a set $ I\subseteq \mathcal{R}$ a \emph{(polynomial) ideal} if
	$p_1,p_2\in I$ implies $p_1+p_2\in I$ and
	$p_1\in I,p_2\in\mathcal{R}$ implies $p_1p_2\in I$.
\end{definition}
\begin{definition}
	We denote by $\langle H^1,\ldots,H^r\rangle$ the smallest (polynomial) ideal containing $\{H^1,\ldots,H^r\}$.
\end{definition}
Given the ideal $\langle H^1,\ldots,H^r\rangle$ and the discrete time derivative $\mathcal{T}\in\mathcal{R}$, we want to determine if $\mathcal{T}\in\langle H^1,\ldots,H^r\rangle$ or if we can write $\mathcal{T}$ in a simpler form, using $\{H^1,\ldots,H^r\}$.
Hence, using multivariate polynomial division, we search for $p^1,\ldots,p^r,p^{r+1}\in\mathcal{R}$ such that
$
	\mathcal{T}=p^1H^1+\ldots+p^rH^r+p^{r+1}.
$
Unfortunately, the resulting remainder $p^{r+1}$ may not be unique \cite[page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although $\mathcal{T}$ belongs to $\langle H^1,\ldots,H^r \rangle$. 
Replacing $\{H^1,\ldots,H^r\}$ by a Gr{\"o}bner basis for the ideal $\langle H^1,\ldots,H^r \rangle$ guarantees the uniqueness of the remainder of the polynomial division. 
Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, a Gr{\"o}bner basis for the polynomial ideal always exists and is finite.
A Gr{\"o}bner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.

@@

	\caption{{\sc FindDiscreteConservedQuantityBasis}}
	\label{alg: findconservedquantitybasis}
\end{algorithm}
We have implemented this algorithm in {\sc FindDiscreteConservedQuantityBasis}.
\begin{example}
In the context of the heat equation from Example \ref{heatex1}, we aim to find a basis for conserved quantities generated by $u$ and $n$ with degree up to 3.
	\begin{EXE}
		LABEL="heateq11"
		(*CODE*)
		variables = <|"indVars"->{n},"depVars"->{u},"rhs"->{u[n]+u[n+1]-2*u[n]+u[n-1]}|>
		FindDiscreteConservedQuantityBasisOperator[variables][<|"degree"->3,"generators"->{u[n],n}|>]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/heateq11ctt.ed.tex}
		\input{executables/heateq11ctt.result.ed.tex}
	\end{small}
The first conserved quantity corresponds to mass conservation and the second to center of mass conservation. 
\end{example}
\section{Examples and applications}
\label{eaa}
We demonstrate our code's capabilities using two examples: the inviscid Burgers equation and a mean-field game system. 
Through these examples, we examine different discretizations, evaluate their conservation properties, and identify conserved quantities.
Although the examples presented feature only first-order time derivatives, the algorithm can handle PDEs with derivatives of any order.

@@

		v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,$ and $

@@

	The related backward-forward system is
		-v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0$ and $

@@

\begin{comment}

@@

\end{comment}
\section{Concluding remarks}
